---
title: "DATA 621 Final"
author: "Adam Gersowitz"
date: "05/13/2021"
output:
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage


```{r, message=FALSE,warning=FALSE, echo=F}
# loading libraries
library(tidyverse)
library(caTools)
library(caret)
library(pROC)
library(knitr)
library(naniar)
library(reshape2)
library(stats)
library(corrplot)
library(e1071)
library(Amelia)
library(jtools)
library(performance)
library(cvms)
library(ROCR)
library(MASS)
library(funModeling)
library(skimr)
library(rnaturalearth)

```




# Data Exploration and Preparation

## Load Input Datasets and Prep


```{r, echo=FALSE,message=FALSE,warning=FALSE}
load_chi2018 <- function() {
  
  df = read.csv("https://raw.githubusercontent.com/davidblumenstiel/CUNY-MSDS-DATA-621/main/Final_Project/chi-2018.csv")
  
  
  x1 = df %>%
    pivot_longer(
      cols = starts_with("wat_"), 
      names_to = "Year",
      names_prefix = "wat_",
      values_to = "water_score"
    ) %>% 
    select("Year", "water_score", "CountryName")
  
  
  x2 = df %>%
    pivot_longer(
      cols = starts_with("san_"), 
      names_to = "Year",
      names_prefix = "san_",
      values_to = "sanitation_score"
    ) %>% 
    select("Year", "sanitation_score", "CountryName")
  
  x3 = df %>%
    pivot_longer(
      cols = starts_with("chmort_"), 
      names_to = "Year",
      names_prefix = "chmort_",
      values_to = "child_mortality"
    ) %>% 
    select("Year", "child_mortality", "CountryName")
  
  
    
  x4 = df %>%
    pivot_longer(
      cols = starts_with("mortality_"), 
      names_to = "Year",
      names_prefix = "mortality_",
      values_to = "mortality_score"
    ) %>% 
    select("Year", "mortality_score", "CountryName")
    
  x5 = df %>%
    pivot_longer(
      cols = starts_with("CHI_v2018_"), 
      names_to = "Year",
      names_prefix = "CHI_v2018_",
      values_to = "CHI_v2018"
    ) %>% 
    select("Year", "CHI_v2018", "CountryName")
  
  out = x1 %>% merge(x2, by = c("CountryName", "Year")) %>%
    merge(x3, by = c("CountryName", "Year")) %>%
    merge(x4, by = c("CountryName", "Year")) %>%
    merge(x5, by = c("CountryName", "Year")) 
}
df <- load_chi2018()

```

## Numerical Summaries

```{r}
summary(df)
```

## Missing Data Check

Missmap Plot illustrates there are missing values for CHI_v2018, child_mortality, mortality_score and water_score. We will remove the records that have missing data.

```{r, message=FALSE, warning=FALSE,echo=FALSE}
missmap(df, main="Missing Values") 
colSums(is.na(df))%>% kable() 



```








## Correlation Plot

Correlation Plot illustrates the relationship between the variables in the input dataset.


```{r,message=FALSE,warning=FALSE,echo=FALSE}

df<-df%>%drop_na()
correlation_table(data=df, target="change_mortality") %>%
  kable() 

```



##  Data Splitting

The Input dataset is split into training and test data using createDataPartition function. The ratio of splitup is 75% for training and 25% for test data.

```{r, warning=FALSE, message=FALSE,echo=FALSE}
set.seed(1005)
training_partition <- createDataPartition(df$change_mortality, p=0.75, list = FALSE, times=1)
train <- df[training_partition, ]
test <- df[-training_partition, ]

train

```


# Build Models

## Stepwise Multiple Linear Regression

```{r, message=FALSE, warning=FALSE,echo=FALSE}

model<- lm(child_mortality~CountryName+sanitation.score.water.score
           , data = train) %>%
  stepAIC(trace = FALSE)
summary(model)


```



