---
title: "DATA 621 HW2"
author: "Adam Gersowitz"
date: "3/14/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE}
library(readr)
library(dplyr)
library(kableExtra)
library(plotly)
library(DescTools)

df <- read_csv("https://raw.githubusercontent.com/agersowitz/DATA-621/main/classification.txt")
df <- data.frame(df)


table(df$class,df$scored.class)%>%
  kbl(caption = "Confusion Matrix") %>%
  kable_classic(full_width = F, html_font = "Cambria")%>%
  kable_styling(latex_options = "HOLD_position")

```

The confusion Matrix above shows the occurrence of True Negatives (119), False Positives (5), False Negatives (30), and True Positives (27).




```{r 3}

classification_metrics <- function(df){
          TP <- nrow(filter(df,class==1 & scored.class==1))
          TN<- nrow(filter(df,class==0 & scored.class==0))
          FP<- nrow(filter(df,class==0 & scored.class==1))
          FN<- nrow(filter(df,class==1 & scored.class==0))
          Acc<- (TP+TN)/(TP+FP+TN+FN)
          CER<- (FP+FN)/(TP+FP+TN+FN)
          pre<- (TP)/(TP+FP)
          sens<-(TP)/(TP+FN)
          spec<-(TN)/(TN+FP)
          f1<-(2*pre*sens)/(pre+sens)
          if (Acc+CER==1){
            stmnt <- paste0("\nSuccess. Accuracy and Classification Error Rate sum to 1.")
          } else {
            stmnt <- paste0("\nFailiure. Accuracy and Classification Error Rate do not sum to 1.")

          }
          return(cat("Classification Metrics",
                     "\nAccuracy = ", round(Acc,3),
                "\nClassification Error Rate = ", round(CER,3),
                stmnt,
                "\nPrecison = ",round(pre,3),
                "\nSensitivity = ",round(sens,3),
                "\nSpecificity = ",round(spec,3),
                "\nF1 Score = ",round(f1,3),"\n\n"
                        
                        
                        ))
}

roc_curve <-function(df){
x<-c()
y<-c()
rx<-0
ry<-0
for (i in 1:nrow(df)){
  yc<-nrow(filter(df, scored.class == 1))
  xc<-nrow(df)
  col<-match("scored.class",names(df))
  if (df[i,col] == 1){
    ry<-ry+1
    rx<-rx
  }else{
      ry<-ry
      rx<-rx+1
    }
  sens<-ry/yc
  specm1<-rx/xc
  x<-append(x,specm1)
  y<-append(y,sens)
  
  dfr<-as.data.frame(cbind(x,y))
  auc<-AUC(x=x,y=y,from = 0, to = 1)
  
  
 
  
}

 plot_ly(dfr, x = ~x, y = ~y, type = 'scatter', mode = 'lines') %>% 
  layout(title = paste0("ROC Curve for Chosen Dataset"),
         annotations = list(text = paste0("Area Under Curve = ",round(auc,3)),  x = .75, y = .25,showarrow=FALSE ),
         xaxis = list(showgrid = FALSE, title = "1-Specificity (false positive rate)"),
         yaxis = list(showgrid = FALSE, title = "Sensitivity (true positive rate)"))
 
}



classification_metrics(df)
roc_curve(df)
```

## 9) Bounds of F1 Score

$$Precision =\frac{TP}{TP+FP}$$
$$Sensitivity=\frac{TP}{TP+FN}$$

$$F1 Score=\frac{2*Precision*Sensitivity}{Precision+Sensitivity}$$

Considering the formulas above we know that Precision and Sensitivity Will be between 0 and 1 (if TP = 0 then they will be 0 and if FN or FP = 0 then they will = 1). So after plugging these minimum and maximum possible values into the F1 Score we will get our range of values.

$$F1 Score=\frac{2*Precision*Sensitivity}{Precision+Sensitivity}=\frac{2*0*0}{0+0}=0$$
$$F1 Score=\frac{2*Precision*Sensitivity}{Precision+Sensitivity}=\frac{2*1*1}{1+1}=\frac{2}{2}=1$$


```{r, message=FALSE}
library(caret)

confusionMatrix(data=as.factor(df$scored.class),reference=as.factor(df$class), positive = "1")
sensitivity(data=as.factor(df$scored.class),reference=as.factor(df$class), positive = "1")
specificity(data=as.factor(df$scored.class),reference=as.factor(df$class), negative = "0")
```

My numbers seem to line up with the built in functions.


```{r, message=FALSE}
library(pROC)

roc1<-roc(df$class,df$scored.probability)
plot(roc1)
```




references : https://acutecaretesting.org/en/articles/roc-curves-what-are-they-and-how-are-they-used#:~:text=An%20ROC%20curve%20shows%20the,TP%2F(TP%2BFN))











